Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench

### 2021-09-16 paper/fips.tex
Update fips.tex and openssl.tex to comform to the rest of the paper's style

\section{SHA3 FIPS Specification in Cryptol}
\label{sec:fips}
Initially it appeared that visual inspection with Cryptol's standard syntax was subpar.
The language's functional folds and list comprehensions do not align easily with the imperative 
and array-based NIST standard. By leveraging Cryptol's versatility, however, we produced a couple 
utility functions to visually mimic the sequential nature of the NIST specification.

@@

By nesting Cryptol's list comprehensions, we easily implemented the pi function in Figure 
\ref{fig:cryptolPi}.

@@

Cryptol declares list comprehension variables after they are used, and the order of the
nested loops is not as apparent. Using a helper function, \emph{for}, we were able to provide
a more clear definition seen in Figure \ref{fig:cryptolAmendedPi}.

@@

Maintaining this state through Cryptol's list comprehensions can be difficult.
The comprehensions require a list to iterate through.Since our \emph{for} utility 
function is simply a wrapper for list comprehensions, it is equally unsuited for 
visual verification on such steps. The condition to stop iterating in the RC function
bounds the number of iterations by [1, 254], which lets us construct such a list to 
iterate through. This leads to a rather ungainly Cryptol definition, however, as 
shown by Figure \ref{fig:cryptolRC}.

@@

Not only is this definition computationally slower as it runs 255 loops every time it is called, 
it also awkwardly indexes into a list of previous states. This differs vastly from the 
NIST specificiation both visually and functionally. Using a new \emph{while} utility function, we 
can declare an initial state, condition function, and body function that will recursively 
be called until the condition is met. It consolodates the information of the previous state 
until the condition is sastified. This definition is foudn in Figure \ref{fig:cryptolamendedRC}.

>>>>

\section{SHA3 FIPS Specification in Cryptol}
\label{sec:fips}
Initially it appeared that visual inspection with Cryptol's standard syntax was subpar.
The language's functional folds and list comprehensions do not align easily with the imperative and array-based NIST standard. 
By leveraging Cryptol's versatility, utility functions can visually mimic the sequential nature of the NIST specification.

@@

By nesting Cryptol's list comprehensions, the pi function can be easily implemented.
This is illustrated in Figure \ref{fig:cryptolPi}.

@@

Cryptol declares list comprehension variables after they are used, and the order of the nested loops is not as apparent. 
Using a helper function, \emph{for}, a clearer definition seen in Figure \ref{fig:cryptolAmendedPi}.

@@

Maintaining this state through Cryptol's list comprehensions can be difficult.
The comprehensions require a list to iterate through. 
Since the \emph{for} utility function is simply a wrapper for list comprehensions, it is equally unsuited for visual verification on such steps. 
The number of iterations in the RC function is bounded by [1, 254] by its modular condition.
This range provides the list to iterate over as required by Cryptol list comprehensions. 
From these observations, a rather ungainly Cryptol definition results.
This definition is shown in Figure \ref{fig:cryptolRC}.

@@

Not only is this definition computationally slower as it runs 255 loops every time it is called, it also awkwardly indexes into a list of previous states. 
This differs vastly from the NIST specificiation both visually and functionally. 
The \emph{while} utility function, accepts an initial state, condition function, and body function to recursively loop. 
It consolodates the information of the previous state until the condition is sastified. 
This definition is found in Figure \ref{fig:cryptolamendedRC}.

### 2021-09-16 paper/openssl.tex
Restructure sections
Update fips.tex and openssl.tex to comform to the rest of the paper's style

\section{Static Iota}\label{sec:iota}
NIST specification invokes the rc method seven times for each iota function call.
As the rc inputs are the same for each round, OpenSSL's optimized implementation of keccak stores 
the output values as a constant table. Initially, we matched out Cryptol execution of the NIST standard 
as closely as possible--including rc method calls in the iota function. We soon found that SAW was unable 
to map our implementation of keccak to OpenSSL's without the use of a constant table.
Our final Cryptol implementation uses the rc definition found in the NIST specification to generate a table 
that is easiy compared to OpenSSL. This allows SAW to generate a proof and permits visual inspection of equivalence.

>>>>

\section{SHA3 OpenSSL in Cryptol}\label{sec:openssl}
The NIST specification invokes the rc method seven times for each iota function call.
As the rc inputs are the same for each round, OpenSSL's optimized implementation of keccak stores the output values as a constant table rather than repeatedly compute them. 
Initially, the Cryptol code to the NIST standard as closely as possible\textemdash including RC method calls in the iota function. 
SAW was unable to map such an implementation of keccak to OpenSSL's implementation without the use of a constant table. 
The final Cryptol implementation uses the RC definition found in the NIST specification to generate a table that is easiy compared to OpenSSL. 
This allows SAW to generate a proof and permits visual inspection of equivalence.
